\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-10-18 eg6-letters/NIM/EG6_RTPC_NIM.tex
more answers and implementations of the comments on the NIM paper

%\begin{keywords}
%\textcolor{red}{Raphael please add the list you provided originally},
%gas electron multipliers.
%\end{keywords}

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beam axis; it is a 284~mm 
     long, 6~mm diameter Kapton straw with a 27~$\mu$m wall.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~$\mu$m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The average electric field in this region is perpendicular to the beam 
      and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

The GEM technology has been chosen for the flexibility of the GEM foils,
which can be easily used to produce a curved amplification surface. Also, 
GEMs are known to have relatively low spark rate~\cite{Beringer:1900zz}, which 
is important when trying to detect 
highly ionizing slow nuclei that deposit large amount of energy. The GEMs for 
this RTPC are made from a Kapton insulator layer, 300~$\mu$m 
thick, sandwiched between two 5~$\mu$m copper layers\footnote{The GEM foils were produced by 
Tech-Etch, Inc.}. The mesh of each GEM 
layer is chemically etched with 50~$\mu$m diameter holes with double-conical 
shapes as illustrated in Figure~\ref{fig:GEMs}. The potential difference 
applied between the two copper layers of the GEM creates a very strong electric 
field in each hole leading to high ionization and amplification.

@@

At the end of the extraction procedure, the azimuthal difference between the detection pad and 
the ionization point ($\Delta\phi$) was extracted as a function of time.  In 
Figure~\ref{fig:DELTA_PHI_TDC}, we show the resulting data points for one bin, 
where the drift path is easily identified and eventually fitted for 
implementation in our reconstruction code.

@@

\bibitem{Beringer:1900zz} 
   J.~Beringer {\it et al.} [Particle Data Group],
  Review of Particle Physics (RPP),
  Phys.\ Rev.\ D {\bf 86}, 010001 (2012).
  %doi:10.1103/PhysRevD.86.010001

>>>>

%\begin{keywords}
%Time projection chamber, gas electron multipliers, alpha particles, DVCS, 
%Nuclear physics
%\end{keywords}

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target forms the detector central z-axis which is 
     along the beamline ; it is a 284~mm long, 6~mm diameter Kapton straw with 
     a 27~$\mu$m wall.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~$\mu$m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The average electric field in this region is perpendicular to the beam 
      and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

The GEM technology has been chosen for the flexibility of the GEM foils,
which can be easily used to produce a curved amplification surface. Also, GEMs 
are known to have relatively low spark rate~\cite{Bachmann:2000kj}, which is 
important when trying to detect highly ionizing slow nuclei that deposit large 
amount of energy. The GEMs for this RTPC are made from a Kapton insulator 
layer, 50~$\mu$m thick, sandwiched between two 5~$\mu$m copper 
layers\footnote{The GEM foils were produced by Tech-Etch, Inc.}. The mesh of 
each GEM layer is chemically etched with 50~$\mu$m diameter holes with 
double-conical shapes as illustrated in Figure~\ref{fig:GEMs}. The potential 
difference applied between the two copper layers of the GEM creates a very 
strong electric field in each hole leading to high ionization and 
amplification.

@@

At the end of the extraction procedure, the azimuthal difference between the 
detection pad and the ionization point ($\Delta\phi$) was extracted as a 
function of time. In Figure~\ref{fig:DELTA_PHI_TDC}, we show the resulting data 
points for one bin in z-coordinate of the RTPC, where the drift path is easily 
identified and eventually fitted for implementation in our reconstruction code.

@@

\bibitem{Bachmann:2000kj} S.~Bachmann {\it et al.},
 Performance of GEM detectors in high intensity particle beams,
 Nucl.\ Instrum.\ Meth.\ A {\bf 470}, 548 (2001).
 %doi:10.1016/S0168-9002(01)00802-6

### 2017-10-18 eg6-letters/NIM/acknowledgment.tex
updated the acknowledgments of the NIM paper
more answers and implementations of the comments on the NIM paper

The authors thank the staff of the Accelerator and Physics Divisions at the 
Thomas Jefferson National Accelerator Facility who made this work 
possible. This work was supported in part by the French Centre National de la 
Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) 
and the U.S. Department of Energy. The 
author (M.~Hattawy) also acknowledges the support of the {\it Consulat 
G\'en\'eral de France \`a J\'erusalem}. 
The Southeastern Universities Research Association operates the Thomas 
Jefferson National Accelerator Facility for the United States Department of 
Energy under contract DE-AC05-06OR23177.

>>>>

The authors thank the staff of the Accelerator and Physics Divisions at the 
Thomas Jefferson National Accelerator Facility who made this work possible.  
This work was supported in part by the French Centre National de la Recherche 
Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) 
and the U.S. Department of Energy. M.~Hattawy also acknowledges the support of 
the {\it Consulat G\'en\'eral de France \`a J\'erusalem}.  The Southeastern 
Universities Research Association operates the Thomas Jefferson National 
Accelerator Facility for the United States Department of Energy under contract 
DE-AC05-06OR23177. This material is based upon work supported by the U.S.  
Department of Energy, Office of Science, Office of Nuclear Physics, under 
contract number DE-AC02-06CH11357.

### 2017-10-18 eg6-letters/NIM/review-1/review-1.tex
more updates to the comments on the nim paper
more answers and implementations of the comments on the NIM paper

\subsection*{Chapter 2}
\item As already said I would not speak of cylindrical GEMs but 
   semi-cylindrical. Moreover, please add the reference to Sauli's paper [11] 
   when speaking the first time about GEM in the text.\\
\textcolor{blue}{ }

\item In my opinion, this chapter should highlight better which are the 
   improvements w.r.t. BoNuS detector to underline the originality of this 
   work. For example, better acceptance? Higher gain? Higher electric 
   stability? Lower material budget? Please quote some old and new values, 
   maybe in a table, to make them evident.\\
\textcolor{blue}{ }

\item Relevant length and thicknesses of the various elements are reported in 
   the text and in the figure, can you please add also the z position of the 
   target w.r.t the detector? To give the complete information, I would also 
   add the value of the induction gap electric field, which is the only missing 
   field value and the value of the reached amplification gain with the three 
   GEM.\\
\textcolor{blue}{ }

\item When speaking of low spark rate the reference [12] is quoted, which is 
   the PDG. Maybe it is better to quote more specific studies like the ones by 
   Bachmann et al.\\
\textcolor{blue}{ }

\item line 119: is the Kapton of the GEM foil really 300 mum thick? It is 
   usually very thin, like 50 mum, in order to have low material budget and 
   very high fields.\\
\textcolor{blue}{ }

\subsection*{ Chapter 3}
\item The description of the electronics looks to me a little not well 
   organized, like a collection of info from previous descriptions from ALICE 
   and BoNuS. It is a series of details, some of which not really relevant for 
   the good understanding of the work described in the paper. So, I would put 
   only some general info (the number of channels/pad, the number of time 
   samples and their width for each channel and the number of  ADC samples and 
   their width for each channel) with the reference where to find more material 
   and just write explicitly the novelties, such as the fact that the 
   acquisition rate was increased (footnote 2).\\
\textcolor{blue}{ }

\item Also, I understand that fig. 5 comes from reference [14], updated with 
   the values of the RTPC instead of the ALICE-TPC, but it seems to me that 
   there are some parts of the picture which have to be fixed. I list them 
   here:\\
- why does it say 768 pads?\\
- why 7.6 microsec?\\
- the anode wire e grid make no sense here\\
- why L1 and L2 (trigger levels?) \\
\textcolor{blue}{ }

\subsection*{ Chapter 4}
	/paragraph 4.1
\item I would really prefer to call it drift "velocity" instead of drift 
   "speed"\\
\textcolor{blue}{ }

\item I don't really understand the procedure to find the drift velocity 
   described here and fig. 7. What is $Nb_{max}$ e $Nb_{max/2}$? Why does 
   $T_{max/2}$ correspond to $Nb_{max/2}$ and why is it used as $T_{max}$? If I 
   make a quick calculation and use $T_{max}$ ~ 65 * 100ns and a drift length 
   of ~ 3cm /cos(23°) ~ 7.5 cm I get a $v_{drift}$ ~ 1.15 cm/microsec << 5 
   cm/microsec which is the average value of drift velocity in drift 
   chambers.\\
The procedure according to me would be:\\
- for each hit associated to a track, whose time information is available, take 
the drift time $t_{drift_{i}}$;\\
- plot the distribution of these $t_{drift_{i}}$;\\
- find the $T_{min}$ = minimum $t_{drift_{i}}$ and $T_{max}$ = maximum 
$t_{drift_{i}}$\\
- compute $delta_{t} = T_{max} - T_{min}$ and extract the drift velocity by 
dividing the track length / $delta_{t}$. Figure 8 would have the $delta_{t}$ 
instead of the $T_{max}$ vs z.\\
Please provide more details on the applied procedure, since I have the open 
questions I wrote before and I could not understand it properly. Moreover, 
please quote the value of the found drift velocity. Is it in accordance with 
the simulations?\\
\textcolor{blue}{ }

\item Figure 6: why are $T_{min}$ and $T_{max}$ not at the drift gap 
   extremities?\\
\textcolor{blue}{ }

@@

\item line 306: "in one bin" should be explained better. I understood it is the 
   subdivision in bins of the z coordinate, but it was not so obvious at the 
   first reading\\
\textcolor{blue}{ }

\item Has the stability of the drift path been counterchecked with simulations 
   varying the field according to the cathode radius variation? How much is the 
   field variation if the radius have some mm of uncertainty?\\
\textcolor{blue}{ }

@@

\subsection*{ Chapter 5}
        /paragraph 5.1	\item Please detail a little bit more the procedure to 
           remove the oscillatory noise. It says "event by event and channel by 
           channel": if the ADC threshold is passed then a noise level is 
           subtracted? If so, how is it calculated?\\
\textcolor{blue}{ }

/paragraph 5.2
\item How was "10.5 mm" cut chosen? is it in xy plane or in 3-dimensions?\\
\textcolor{blue}{ }

\item How was the "10 hits" cut chosen? What is the mean number of primary 
   ionization in this gas mixture?\\
\textcolor{blue}{ }

@@

        /paragraph 6.2
\item I understand that efficiency is a ratio between two integrals, each 
   integral being the area of the Gaussian which fits the recoil mass 
   distribution in the inclusive and exclusive case. If so, would it be 
   possible to please draw both fits on the figure 14 and provide the integrals 
   to compute the mean efficiency? If I misunderstood, please comment.\\
\textcolor{blue}{ }

\subsection*{ Chapter 7} \item Is it "high rate environment" or "high readout 
   rate"? They are connected, but what is the rate the RTPC underwent during 
   data taking?\\
\textcolor{blue}{ }

\item I understand the resolution and efficiency have been evaluated with the 2 
   GeV electrons. Is it foreseen to analyze also the 6 GeV electron runs, which 
   were the actual data taking? Is there an evaluation of the foreseen RTPC 
   performances in that case?\\
\textcolor{blue}{ }

>>>>

\subsection*{Chapter 2}
\item As already said I would not speak of cylindrical GEMs but 
   semi-cylindrical. Moreover, please add the reference to Sauli's paper [11] 
   when speaking the first time about GEM in the text.\\
\textcolor{blue}{cylinderical GEMs-> semi-cylinderical. The reference has been added.}

\item In my opinion, this chapter should highlight better which are the 
   improvements w.r.t. BoNuS detector to underline the originality of this 
   work. For example, better acceptance? Higher gain? Higher electric 
   stability? Lower material budget? Please quote some old and new values, 
   maybe in a table, to make them evident.\\
\textcolor{red}{To the group: Do we have such a comparison? }

\item Relevant length and thicknesses of the various elements are reported in 
   the text and in the figure, can you please add also the z position of the 
   target w.r.t the detector? To give the complete information, I would also 
   add the value of the induction gap electric field, which is the only missing 
   field value and the value of the reached amplification gain with the three 
   GEM.\\
\textcolor{blue}{- The target tube is the central z-axis of the detector. The 
text is modified to make it clean.\\}
\textcolor{red}{- Regarding the electric field, it was 2.8 kV/cm in BoNuS, does 
anyone know our exact field in this gap?\\}
\textcolor{blue}{- The gain of such GEM configuration has been studied by 
   CERN's Gas Detector Development Group (Fabio Sauli, Progress with the Gas 
   Electron Multiplier, 2nd Workshop on Advanced Transition Radiation Detectors 
   for Accelerator and Space Applications (Bari, Sept. 4-7, 2003), Nucl. Instr. 
   and Meth. A522(2004)93). The results, see the reference, demonstrate that 
one should not simply add gains to each other for multilayer GEM 
amplifications.  We did not evaluate precisely the overall gain in our chamber, 
so we do not have a value for our RTPC GEM system at this point.}

\item When speaking of low spark rate the reference [12] is quoted, which is 
   the PDG. Maybe it is better to quote more specific studies like the ones by 
   Bachmann et al.\\
\textcolor{blue}{Thanks. It has been replaced.}

\item line 119: is the Kapton of the GEM foil really 300 mum thick? It is 
   usually very thin, like 50 mum, in order to have low material budget and 
   very high fields.\\
\textcolor{blue}{Sorry typo. It is 50 $mu$m}

\subsection*{ Chapter 3}
\item The description of the electronics looks to me a little not well 
   organized, like a collection of info from previous descriptions from ALICE 
   and BoNuS. It is a series of details, some of which not really relevant for 
   the good understanding of the work described in the paper. So, I would put 
   only some general info (the number of channels/pad, the number of time 
   samples and their width for each channel and the number of  ADC samples and 
   their width for each channel) with the reference where to find more material 
   and just write explicitly the novelties, such as the fact that the 
   acquisition rate was increased (footnote 2).\\
\textcolor{red}{Raphael, Can you check here?}

\item Also, I understand that fig. 5 comes from reference [14], updated with 
   the values of the RTPC instead of the ALICE-TPC, but it seems to me that 
   there are some parts of the picture which have to be fixed. I list them 
   here:\\
- why does it say 768 pads?\\
- why 7.6 microsec?\\
- the anode wire e grid make no sense here\\
- why L1 and L2 (trigger levels?) \\
\textcolor{red}{Raphael, Can you check here? }

\subsection*{ Chapter 4}
	/paragraph 4.1
\item I would really prefer to call it drift "velocity" instead of drift 
   "speed"\\
\textcolor{red}{To the group: We do have radial and azimuthal drift velocities, 
but we measure the average of their sum. }

\item I don't really understand the procedure to find the drift velocity 
   described here and fig. 7. What is $Nb_{max}$ and $Nb_{max/2}$? Why does 
   $T_{max/2}$ correspond to $Nb_{max/2}$ and why is it used as $T_{max}$? If I 
   make a quick calculation and use $T_{max}$ ~ 65 * 100ns and a drift length 
   of ~ 3cm /cos(23°) ~ 7.5 cm I get a $v_{drift}$ ~ 1.15 cm/microsec << 5 
   cm/microsec which is the average value of drift velocity in drift 
   chambers.\\
The procedure according to me would be:\\
- for each hit associated to a track, whose time information is available, take 
the drift time $t_{drift_{i}}$;\\
- plot the distribution of these $t_{drift_{i}}$;\\
- find the $T_{min}$ = minimum $t_{drift_{i}}$ and $T_{max}$ = maximum 
$t_{drift_{i}}$\\
- compute $delta_{t} = T_{max} - T_{min}$ and extract the drift velocity by 
dividing the track length / $delta_{t}$. Figure 8 would have the $delta_{t}$ 
instead of the $T_{max}$ vs z.\\
Please provide more details on the applied procedure, since I have the open 
questions I wrote before and I could not understand it properly. Moreover, 
please quote the value of the found drift velocity. Is it in accordance with 
the simulations?\\
\textcolor{blue}{ }

\item Figure 6: why are $T_{min}$ and $T_{max}$ not at the drift gap 
   extremities?\\
   \textcolor{blue}{This is a schematic to illustrate the definitions of 
   $T_{min}$ and $T_{max}$ for a chain hits. The figure can be modified if 
needed.}

@@

\item line 306: "in one bin" should be explained better. I understood it is the 
   subdivision in bins of the z coordinate, but it was not so obvious at the 
   first reading\\
\textcolor{blue}{Cleared}

\item Has the stability of the drift path been counter checked with simulations 
   varying the field according to the cathode radius variation? How much is the 
   field variation if the radius have some mm of uncertainty?\\
\textcolor{blue}{We did not study this case.}

@@

\subsection*{ Chapter 5}
        /paragraph 5.1

\item Please detail a little bit more the procedure to remove the oscillatory 
   noise. It says "event by event and channel by channel": if the ADC threshold 
   is passed then a noise level is subtracted? If so, how is it calculated?\\
\textcolor{red}{Nathan, Can you work on it. }

/paragraph 5.2
\item How was "10.5 mm" cut chosen? is it in xy plane or in 3-dimensions?\\
\textcolor{blue}{The distance between hits are calculated in 3D. This cut has 
been chosen because it enables us to jump longer gaps from dead areas of the 
readout pads without increasing the level of spurious tracks.}

\item How was the "10 hits" cut chosen? What is the mean number of primary 
   ionization in this gas mixture?\\
   \textcolor{blue}{Figure \ref{fig:nb_hits} shows the number of hits 
   distribution for the collected elastic $^4$He tracks from the 1.2 GeV 
dataset. The mean number of hits is around 135 and the 10 hits cut at the 
reconstruction level is a loose cut to reduce the level of the spurious 
tracks.}

\begin{figure}[h!]
   \centering
   \includegraphics[scale=0.45]{nb_hits.png}
   \caption{the number of hits distribution for the collected elastic $^4$He 
   tracks from the 1.2 GeV dataset.} \label{fig:nb_hits}
\end{figure}

@@

        /paragraph 6.2
\item I understand that efficiency is a ratio between two integrals, each 
   integral being the area of the Gaussian which fits the recoil mass 
   distribution in the inclusive and exclusive case. If so, would it be 
   possible to please draw both fits on the figure 14 and provide the integrals 
   to compute the mean efficiency? If I misunderstood, please comment.\\
\textcolor{red}{Nathan, do you still have the individual integrals? }

\subsection*{ Chapter 7}

\item Is it "high rate environment" or "high readout rate"? They are connected, 
   but what is the rate the RTPC underwent during data taking?\\
\textcolor{red}{Can we find such information in the logbook?}

\item I understand the resolution and efficiency have been evaluated with the 2 
   GeV electrons. Is it foreseen to analyze also the 6 GeV electron runs, which 
   were the actual data taking? Is there an evaluation of the foreseen RTPC 
   performances in that case?\\
\textcolor{blue}{Typically, the elastic reaction is used for such performance 
evaluation. But the elastic cross sections is really small at 6 GeV beam 
energy.}
